The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-02-08 src/ms.tex
First draft of funding section

Summarize major pieces of development since the last paper (since v2.0). A few
ideas (taken from What's New pages) below, but feel free to expand on or remove
from this list:
\begin{itemize}
    \item Support for representing and transforming velocity data in coordinates, epoch propagation (v3.0)
    \item Improved support for astronomical time series: TimeSeries object (v3.2), Box Least Squares periodogram (v3.1)
    \item Overall improved support of Quantity throughout numpy (v4.0) and scipy
    \item Native support for Time, Quantity, and SkyCoord objects in Astropy tables
    \item APE 14 WCS
    \item A new SpectralCoord class for representing and transforming spectral quantities
    \item Support for dask/large arrays in several places in the core package (4.1, 5.0, and maybe other versions)
\end{itemize}

@@

Summarize funding sources (Moore, NASA) and amounts and what this has been used
for.

>>>>

Summarize major pieces of development since the last paper (since v2.0). A few
ideas (taken from What's New pages) below, but feel free to expand on or remove
from this list:
\begin{itemize}
    \item Support for representing and transforming velocity data in coordinates, epoch propagation (v3.0)
    \item Improved support for astronomical time series: TimeSeries object (v3.2), Box Least Squares periodogram (v3.1)
    \item Overall improved support of Quantity throughout numpy (v4.0) and scipy
    \item Native support for Time, Quantity, and SkyCoord objects in Astropy tables
    \item TODO: something about WCS, SpectralCoord?
\end{itemize}

@@

% Summarize funding sources (Moore, NASA) and amounts and what this has been
% used for.

With the goal of establishing a sustainable financial model for the Astropy
project, the Coordination Committee under the new governance charter (see \S
\ref{sec:project-governance}) established a standing Finance committee.
This committee overlooks the planning and disbursement of finances on behalf
of the project. The Astropy project accepts funds from institutions as well as
individuals through NUMFOCUS\footnote{NumFOCUS is a 501(c)(3) nonprofit that
supports and promotes world-class, innovative, open source scientific computing.},
the fiscal sponsor of the project. Previously, no direct financial support
was available for the project development, and NUMFOCUS covered the incurred
operational costs. However, transitioning to long-term financial stability meant
having an influx of funds in the form of institutional support \citep{fundmodel},
and hence the Astropy project applied for and successfully acquired several grants.
The two major ones are briefly summarized below.

\begin{itemize}
    \item Moore: The Gordon and Betty Moore Foundation awarded the Astropy
    project a \$900,000 grant in 2019 for the proposal ``Sustaining and Growing
    the Astropy Project". Over a three-year term, this grant supported sustenance
    of the project by providing funds for (1) the development and maintenance of
    the \astropypkg and its infrastructure to existing project members and
    targeted hires, (2) the transition of long-term users to contributors and
    maintainers through mentorship efforts, (3) the formalization of a governance
    structure (see \S \ref{sec:project-governance}), and (4) the improvement
    of equity, diversity, and inclusion efforts in the project (see \S
    \ref{sec:project-ide}). Some of the money supported travel for meetings,
    conferences, and workshops. This generous grant helped reduce reliance on
    volunteer-driven maintenance of the openly-developed package and paved
    the way to acquire funding from federal agencies: a big milestone for the
    success of the project.

    \item NASA: A successful proposal to the NASA ROSES-2020 call, section E.7
    (Support for Open Source Tools, Frameworks, and Libraries), provided the
    Astropy project \$634,590 to support its work. Funded work for a term of
    three years includes (1) project-wide infrastructure maintenance and
    improvements, (2) targeted work on areas of the \astropy core and
    coordinated packages that need support, (3) enhancements to the Learn
    Astropy ecosystem and educational materials (see \S \ref{sec:learn}),
    and (4) support for the \astropy affiliated packages. These goals
    support the sustenance of the project by encouraging further
    development of the package ecosystem and community engagement.
\end{itemize}

In addition to the two grants described above, the project benefited from
a Gemini Observatory contract under its Science User Support Department,
which funded development work to support both the Astropy and DRAGONS
projects. The project also received a total of CAD 31,023 from the Dunlap
Seed Funding program at the University of Toronto to develop educational
resources for scientific software packaging within the Learn Astropy
framework. All these successful proposals show that the Astropy project
is vital for the astronomical infrastructure and community at large, and
strengthen its promising and sustainable future.
